---
title: "Comparing the Effect of Vitamin C and Orange Juice on Guinea Pig Tooth Length"
author: "Jason Dyer"
date: "April 19, 2019"
output: pdf_document
---

###Overview
This report will analyze on the effect of vitamin c and orange juice on the length of teeth in guinea pigs.  The first part will be a basic summary of the data.  The second part will test several hypothesis on the effect of vitamin C and orange juice has on teeth length.

###Load the Data and Libraries
```{r echo=TRUE, warning=FALSE, results="hide", message=FALSE}
library(datasets)
library(pastecs)
library(ggplot2)
data(ToothGrowth)
```

###Data Summary
```{r}
stat.desc(ToothGrowth$len, basic=TRUE, desc=TRUE, norm=TRUE, p=0.95)
```

Our dataset contains 60 observations with a mean of 18.81, a median of 19.25, a minimum value of 4.2 and a maximum value of 33.9.  The negative skewness and kurtosis results indicate that the data is slightly skewed negatively and has a flat distribution.  The Shaprio-Wilk test results show that the data is normally distributed.  Below is a histogram of the dataset:

```{r echo = FALSE}
hist(ToothGrowth$len,
     main = "Histogram of Teeth Length",
     xlab = "Teeth Length",
     col = "green")
```

Plotting teeth length by dosage:
```{r echo = FALSE}
ggplot(ToothGrowth,aes(factor(dose),len,fill=supp)) +
    geom_bar(stat="identity", position = "dodge")+
    xlab("Dose (mg)") +
    ylab("Teeth Length") +
    labs(title = "Teeth Length by Dose and Supplement")+
    theme(plot.title = element_text(hjust=0.5))
```

The plot suggests that orange juice contributes to longer teeth than vitamin C at a vitamin C dosage of 0.5 mg/day and 1 mg/day but at 2 mg/day, the effects of orange juice and vitamin C are similar.

### Use confidence intervals and hypothesis tests to compare tooth growth by supplement and dose.
##### Hypothesis 1
There is no difference on teeth length between orange juice and vitamin C regardless of vitamin C dosage.
```{r}
h1<-t.test(len~supp, data=ToothGrowth)
h1$conf.int
h1$p.value
```
The confidence interval includes zero and has a p-value above 5%.  Therefore, we fail to reject the null hypothesis.

##### Hypothesis 2
There is no difference on teeth length between orange juice and vitamin C dosage of 0.5 mg/day.
```{r}
h2<-t.test(len~supp, data= subset(ToothGrowth,dose==0.5))
h2$conf.int
h2$p.value
```
The confidence level does not include 0 and has a p-value below 5%.  We can reject the null hypothesis and accept the hypothesis that orange juice does increase teeth length more than 0.5 mg/day of vitamin C.

##### Hypothesis 3
There is no difference on teeth length between orange juice and vitamin C dosage of 1.0 mg/day.
```{r}
h3<-t.test(len~supp, data= subset(ToothGrowth,dose==1))
h3$conf.int
h3$p.value
```
The confidence level does not include 0 and has a p-value below 5%.  We can reject the null hypothesis and accept the hypothesis that orange juice does increase teeth length more than 1.0 mg/day of vitamin C.

##### Hypothesis 4
There is no difference on teeth length between orange juice and vitamin C dosage of 2.0 mg/day.
```{r}
h4<-t.test(len~supp, data= subset(ToothGrowth,dose==2))
h4$conf.int
h4$p.value
```
The confidence interval includes zero and has a p-value above 5%.  Therefore, we fail to reject the null hypothesis.

###Conclusion
Based on the results of the hypothesis test, vitamin C is no better than orange in the overall data set but at a vitamin C dosage of 0.5 mg/day and 1.0 mg/ day orange juice will provide more tooth than growth than vitamin C.

###Appendix
Below is the code necessary for the displayed plots

Histogram
```{r results= FALSE}
hist(ToothGrowth$len,
     main = "Histogram of Teeth Length",
     xlab = "Teeth Length",
     col = "green")
```

Plotting teeth length by dosage:
```{r results= FALSE}
ggplot(ToothGrowth,aes(factor(dose),len,fill=supp)) +
    geom_bar(stat="identity", position = "dodge")+
    xlab("Dose (mg)") +
    ylab("Teeth Length") +
    labs(title = "Teeth Length by Dose and Supplement")+
    theme(plot.title = element_text(hjust=0.5))
```
