---
title: "Comparing Exponential Distribution and the Central Limit Theorem"
author: "Jason Dyer"
date: "April 19, 2019"
output: pdf_document
---

```{r echo=FALSE, message=FALSE,warning=FALSE}
#Load Libraries
library(ggplot2)
```

```{r echo=FALSE, message=FALSE,warning=FALSE}
# Variables
lambda<-0.2
n<-40 # number of exponentials
sims<-1000 # number of simulations

#set seed
set.seed(3)
```

###Synopsis
This project will investigate the exponential distribution and compare it with the Central Limit Theorem.  The rate (lambda) will be 0.2 for all of the simulations.  The analysis will compare the distribution of averages and variance of 40 exponentials over a thousand simulations.

##Simulation
I created a matrix of the exponential distribution at a rate of 0.2 for 40 exponentials over 1,000 simulations. Then the mean from the simulations was calculated.    

```{r echo=FALSE, message=FALSE,warning=FALSE}
#Create Matrix
ed<-matrix(rexp(n*sims,lambda),sims)
```

```{r echo=FALSE, message=FALSE,warning=FALSE}
#Row Means
matrixmean<-rowMeans(ed)
```

```{r echo=FALSE, message=FALSE,warning=FALSE}
#Theoritcal Mean
theoryMean<- 1/lambda

#Sample Mean
sampleMean<- mean(matrixmean)
```

###Mean Comparison
There is very little difference between the theoretical mean and the sample mean:
```{r message=FALSE, warning=FALSE}
cat("Theory Mean:",theoryMean,"Sample Mean:",sampleMean)
```

Visually, the theoretical mean and the sample means virtually overlap at 5
```{r echo=FALSE,message==FALSE,warning=FALSE}
hist(matrixmean, xlab= "Sample Mean", main = "Theoretical vs Sample Means")
    abline(v=theoryMean, lty=5, col="black")
    abline(v=sampleMean, col="red")
```

###Variance Comparison
```{r echo=FALSE,message=FALSE,warning=FALSE}
##Compare Variance
    
theoryVariance<-(1/lambda)^2/n
sampleVariance<-var(matrixmean)
```

Similiarlly, we can see that there very little difference between the theoretical variance and the sample variance:

```{r message=FALSE,warning=FALSE}
cat("Theory Variance:",theoryVariance,"Sample Variance:",sampleVariance)
```

###Compare Distribution
Graphing the exponential data shows that its distribution is approximately normal and almost exactly matches with the theoretical distribution.
```{r echo=FALSE, message=FALSE,warning=FALSE}
meanDF<-data.frame(matrixmean)
ggplot(data=meanDF, aes(x=matrixmean)) +
    geom_histogram(binwidth=0.1,aes(y=..density..),alpha=0.2)+
    stat_function(fun=dnorm,args=list(mean=theoryMean),aes(color="Theoretical"),lty=5) +
    stat_function(fun=dnorm,args=list(mean=sampleMean),aes(color="Sample"))+
    scale_colour_manual(values=c("red","black"))+
    theme(panel.grid.major = element_blank(),panel.grid.minor=element_blank())+
    labs(x= "Mean", y="Density",colour="Legend")+
    ggtitle("Theoretical vs Sample Distribution")+
    theme(plot.title = element_text(hjust=0.5))
```

##Conclusion
The Central Limit Theorem tells us that given a sufficiently large sample size from a population the mean of all samples will be approximately equal to the mean of the population and that the distribution of the sample means will be approximately normal.  The analysis has shown that our sample mean is close to the population mean and that the sample mean is approximately normal as predicted by the Central Limit Theorem.

##Appendix
Below is the code for the mean and variance calculations as well as the plots

```{r results=FALSE, message=FALSE,warning=FALSE}
#Load Libraries
library(ggplot2)
```

```{r results=FALSE, message=FALSE,warning=FALSE}
# Variables
lambda<-0.2
n<-40 # number of exponentials
sims<-1000 # number of simulations

#set seed
set.seed(3)
```

```{r results=FALSE, message=FALSE,warning=FALSE}
#Create Matrix
ed<-matrix(rexp(n*sims,lambda),sims)
```

```{r results=FALSE, message=FALSE,warning=FALSE}
#Row Means
matrixmean<-rowMeans(ed)
```

```{r results=FALSE, message=FALSE,warning=FALSE}
#Theoretical Mean
theoryMean<- 1/lambda

#Sample Mean
sampleMean<- mean(matrixmean)
```

```{r results=FALSE,message==FALSE,warning=FALSE}
#Mean Compare Histogram
hist(matrixmean, xlab= "Sample Mean", main = "Theoretical vs Sample Means")
    abline(v=theoryMean, lty=5, col="black")
    abline(v=sampleMean, col="red")
```

```{r results=FALSE,message=FALSE,warning=FALSE}
##Compare Variance
theoryVariance<-(1/lambda)^2/n
sampleVariance<-var(matrixmean)
```

```{r results=FALSE, message=FALSE,warning=FALSE}
#Theoretical vs Sample Distribution
meanDF<-data.frame(matrixmean)
ggplot(data=meanDF, aes(x=matrixmean)) +
    geom_histogram(binwidth=0.1,aes(y=..density..),alpha=0.2)+
    stat_function(fun=dnorm,args=list(mean=theoryMean),aes(color="Theoretical"),lty=5) +
    stat_function(fun=dnorm,args=list(mean=sampleMean),aes(color="Sample"))+
    scale_colour_manual(values=c("red","black"))+
    theme(panel.grid.major = element_blank(),panel.grid.minor=element_blank())+
    labs(x= "Mean", y="Density",colour="Legend")+
    ggtitle("Theoretical vs Sample Distribution")+
    theme(plot.title = element_text(hjust=0.5))
```